\Large \bf DRAFT: Direct-style process creation on Linux

### 2020-05-27 research/paper/paper.tex
paper stufff

\begin{document}
\newcommand{\twosigma}{{[tech company]}}
\newcommand{\githuburl}{[Github url]}
\date{}
\title{\Large \bf DRAFT: Direct-style process creation on Linux}
% \author{
% {\rm Spencer Baugh}\\
% Two Sigma
% }
\maketitle
\begin{abstract}
Traditional process creation interfaces,
such as fork and spawn,
are complex to use, limited in power, and difficult to abstract over.
We develop a new process creation interface for Linux
which allows a program to create a child process in a non-running state
and initialize the new process by operating on it from the outside.
This method of process creation results in more comprehensible programs, 
has better error handling,
is comparably efficient,
and is more amenable to abstraction.
Our implementation is immediately deployable without kernel modifications on any recent Linux kernel version.
\end{abstract}

@@

Our contribution in this paper is such an API, in the form of the \texttt{rsyscall} library.
\texttt{rsyscall} is a
language-specific, object-capability-model, low-abstraction library for the Linux system call interface,
bypassing libc.\cite{capmyths}\cite{robust_composition}
The \texttt{rsyscall} project is currently focused on supporting Python,
but the library can be ported to other languages.
The examples we show in this paper will be in Python,
but generalize easily.
\texttt{rsyscall} is open source and available from
\githuburl{}.

@@

Direct-style \texttt{clone} has acceptable performance cost,
and can outperform \texttt{fork} in some cases.
Like a spawn-style interface,
we can call \texttt{clone}
without worse performance in the presence of large memory usage,
and without the possibility of bugs in the presence of multi-threading.
The performance overhead of creating processes with direct-style clone
is comparable to the performance cost of creating processes while inside a Linux namespace.
At \twosigma{}, we've used direct-style process creation
to implement a production library for distributed system deployment.
We evaluate in further detail in section \ref{evaluation}.
\section{Examples}\label{examples}
In this section,
we'll demonstrate direct-style process creation
by using several sophisticated features of Linux processes.
For concision, we assume that we are running with sufficient privileges,
but these examples can all be run without privileges with appropriate use of \verb|CLONE_NEWUSER|.
\cite{user_namespaces}\cite{unpriv_fuse}

@@

With improved process creation techniques,
these features can be used directly by programmers
instead of by configuring stand-alone software.
In this way,
individual real-world applications can pick and choose the features that are useful for them,
as we describe in section \ref{realworld}.
\subsection{Abstraction through file descriptor passing}\label{fd_abstraction}
In Linux (and Unix in general),
a program can pass file descriptors from the parent process to a child process
by using file descriptor inheritance.\cite{tlpi}
When a process is created, all open file descriptors are copied from the parent process to the child process.
The child process can then make independent use of the file descriptors.
We'll discuss this important feature further in section \ref{fdtables}.

@@

For \texttt{unshare(CLONE\_FILES)}, we perform \texttt{close\_cloexec} automatically in userspace.
For \texttt{clone}, we perform \verb|close_cloexec| only on user request,
so that the user can control when \verb|inherit_fd| stops working.
Even if the user never requests that unused file descriptors be closed,
as long as they call \texttt{execve}, the table will be cleaned at that point.
In both of these cases,
we would benefit from a new system call which takes a list of file descriptors
and closes all \texttt{CLOEXEC} file descriptors except for those on the list;
this would replace our userspace implementation of \verb|close_cloexec|.

@@

As a result, though implementations in native code and in the kernel would likely remove most of this overhead,
we have chosen to not invest effort into optimizing process creation at this micro-level,
to preserve implementation simplicity;
such optimization is reserved for future work.
\subsection{Nested processes, \texttt{CLONE} flags}\label{clone_bench}
\begin{figure}
\centering
 \includegraphics[width=0.5\textwidth]{clone_bench}
 \caption{Time to create different processes from \texttt{rsyscall}, by variety of parent process and child process}
 \label{fig:clone_bench}
\end{figure}
As demonstrated in sections \ref{pidns} and \ref{shared_fd_table},
\texttt{rsyscall} supports nested operation.
That is, we can spawn a child from one of our children;
this can impact performance.
\texttt{clone} also supports specifying a variety of flags to create new namespaces when creating the child,
which can also impact performance.

>>>>

\begin{document}
\newcommand{\twosigma}{{[tech company]}}
\newcommand{\githuburl}{[Github url removed]}
\date{}
\title{\Large \bf DRAFT: Direct-style process creation on Linux}
% \author{
% {\rm Spencer Baugh}\\
% Two Sigma
% }
\maketitle
\begin{abstract}
Traditional process creation interfaces,
such as \texttt{fork} and \texttt{spawn},
are complex to use, limited in power, and difficult to abstract over.
We develop a new process creation interface for Linux
which allows a program to create a child process in a non-running state
and initialize the new process by operating on it from the outside.
This method of process creation results in more comprehensible programs, 
has better error handling,
is comparably efficient,
and is more amenable to abstraction.
Our implementation is immediately deployable without kernel modifications on any recent Linux kernel version.
\end{abstract}

@@

Our main contribution in this paper is such an API, in the form of the \texttt{rsyscall} library.
\texttt{rsyscall} is a
language-specific, object-capability-model, low-abstraction library for the Linux system call interface,
bypassing libc.\cite{capmyths}\cite{robust_composition}
The \texttt{rsyscall} project is currently focused on supporting Python,
but the library can be ported to other languages.
The examples we show in this paper will be in Python,
but generalize easily.
\texttt{rsyscall} is open source and available from
\githuburl{}.

@@

Direct-style \texttt{clone} has acceptable performance cost,
and can outperform \texttt{fork} in some cases.
Like a spawn-style interface,
we can call \texttt{clone}
without worse performance in the presence of large memory usage,
and without the possibility of bugs in the presence of multi-threading.
The performance overhead of creating processes with direct-style clone
is comparable to the performance cost of creating processes while inside a Linux namespace.
At \twosigma{}, we've used direct-style process creation
to implement a production library for distributed system deployment.
We evaluate in further detail in section \ref{evaluation}.

In summary, this paper makes the following contributions:
\begin{compactitem}
\item We coin the term ``direct-style process creation''
  to refer to a previously-unnamed style of process creation which is common on academic operating systems.
\item We built \texttt{rsyscall}, a library for the Linux system call interface following the object-capability model.
\item Using \texttt{rsyscall}, we built the first implementation of direct-style process creation on a Unix-like kernel.
\end{compactitem}
\texttt{rsyscall} is open source and available from \githuburl{}.
\section{Examples}\label{examples}
In this section,
we'll demonstrate direct-style process creation
by using several sophisticated features of Linux processes.
For concision, we assume that we are running with sufficient privileges,
but these examples can all be run without privileges with appropriate use of \verb|CLONE_NEWUSER|.
\cite{user_namespaces}\cite{unpriv_fuse}

@@

With improved process creation techniques,
these features can be used directly by programmers
instead of by configuring stand-alone software.
In this way,
individual real-world applications can pick and choose the features that are useful for them,
as we describe in section \ref{realworld}.

In these examples and in this paper,
we'll have multiple processes,
all of which are under the control of a single program with a single flow of control.
Because of the absence of multiple concurrent control flows,
these examples could be executed interactively by a human operator,
typing them in line by line in a REPL.
To emphasize this point, we use the first person plural --- ``we'' ---
when describing the actions of the single program in these examples and in later sections.
\subsection{Abstraction through file descriptor passing}\label{fd_abstraction}
In Linux (and Unix in general),
a program can pass file descriptors from the parent process to a child process
by using file descriptor inheritance.\cite{tlpi}
When a process is created, all open file descriptors are copied from the parent process to the child process.
The child process can then make independent use of the file descriptors.
We'll discuss this important feature further in section \ref{fdtables}.

@@

When we call \texttt{unshare(CLONE\_FILES)},
we perform \texttt{close\_cloexec} automatically in userspace.
For \texttt{clone}, we perform \verb|close_cloexec| only on user request,
so that the user can control when \verb|inherit_fd| stops working.
Even if the user never requests that unused file descriptors be closed,
as long as they call \texttt{execve}, the table will be cleaned at that point.
In both of these cases,
we would benefit from a new system call which takes a list of file descriptors
and closes all \texttt{CLOEXEC} file descriptors except for those on the list;
this would replace our userspace implementation of \verb|close_cloexec|.

@@

As a result, though implementations in native code and in the kernel would likely remove most of this overhead,
we have chosen to not invest effort into optimizing process creation at this micro-level,
to preserve implementation simplicity;
such optimization is reserved for future work.
\subsubsection{Nested processes, \texttt{CLONE} flags}\label{clone_bench}
\begin{figure}
\centering
 \includegraphics[width=0.5\textwidth]{clone_bench}
 \caption{Time to create different processes from \texttt{rsyscall}, by variety of parent process and child process}
 \label{fig:clone_bench}
\end{figure}
As demonstrated in sections \ref{pidns} and \ref{shared_fd_table},
\texttt{rsyscall} supports nested operation.
That is, we can spawn a child from one of our children;
this can impact performance.
\texttt{clone} also supports specifying a variety of flags to create new namespaces when creating the child,
which can also impact performance.
